Named Entity Recognition (NER), a foundational Natural Language Processing task, extracts entities (persons, organisations, dates) from raw text. Yet real-world applications rarely afford a static class inventory, and the legal domain is an extreme case: new statutes, parties, instruments, courts and jurisdictions appear continually, expert annotation is expensive, and errors carry material consequences. Open-world NER addresses this by assigning types not fixed at training time, but the systems that do so are built and evaluated on general-domain text, so whether they survive a move to specialised legal text is untested.

We study whether OPENER, a frugal and accurate open-world NER pipeline assembled entirely from off-the-shelf pretrained components, transfers to legal subjects without any legal-domain retraining. We reuse its contrastively fine-tuned, Matryoshka-truncatable entity embedder unchanged behind a frozen open-world detector, and benchmark its two typing heads, a label-free zero-shot head (OPENER-ZS) and a supervised linear probe (OPENER-Sup), on four legal NER corpora spanning three languages: English (SEC filings and Indian court judgments), Portuguese (Brazilian court decisions) and German (federal court decisions). Following the general-domain protocol, we report quality (Adjusted Mutual Information), latency and energy jointly, with every OPENER score averaged over three embedder-retraining seeds.

Transfer holds strongly under a pure English domain shift and degrades once a cross-lingual shift is added. OPENER-ZS stays among the most competitive label-free systems on English legal text (up to 49.1 gold AMI), yet its standing slips relative to the general domain, where the same head led every compared zero-shot system: here it falls behind the GLiNER family and, on average, behind OWNER, and weakens sharply cross-lingually (23.8--37.4 gold AMI). OPENER-Sup is barely affected by the domain shift (61.4 mean gold AMI, against 62.3 in the general domain) and posts the best end-to-end AMI throughout (35.9). Its balanced linear probe carves separating hyperplanes fine enough to distinguish closely related legal types, such as lawyers from judges or statutes from case law, even where the shared general-purpose embedding space places their vectors close together, a distinction the zero-shot head's label-name matching cannot sharpen without target supervision. End-to-end, every system compresses into a narrow band set by the English detector's recall (down to 18.7\% on German), at one to two orders of magnitude less latency and energy than the generative baselines. These results indicate that, for legal and other applied domains where labelled examples are available, the supervised head is the more practical operating point, and that in-language detection, not typing, is where further domain adaptation would help first.
