\section{Introduction}
\label{sec:introduction}

Legal text is a demanding setting for named entity recognition (NER). Entity inventories are open and shifting (new statutes, case citations, parties, courts, contractual instruments, jurisdictions), documents are long and formulaic, and annotation requires expert effort, so re-annotating and retraining a closed-set tagger for every new type or sub-domain is rarely practical. This is precisely the regime that \emph{open-world} NER targets: recognising and typing entities whose categories were not fixed at training time. Notable open-world systems take different routes to this goal: GLiNER \citep{zaratiana2024gliner} scores candidate spans against type names supplied as queries, GNER \citep{ding2024gner} recasts NER as text generation, OWNER \citep{genest2025owner} trains a dedicated entity encoder and clusters mentions into dynamically named types, and instruction-tuned language models can be prompted directly, at the cost of far higher inference compute. All are trained and evaluated on general-domain text, so whether an open-world system trained on \emph{general-domain} data can simply be \emph{reused} on legal text, without any legal-domain retraining, is left open. We study this with OPENER \citep{anon2026opener}, a frugal open-world pipeline assembled from off-the-shelf pretrained components (a frozen detector and a contrastively fine-tuned Matryoshka embedder), which on a general-domain benchmark is competitive in accuracy while staying one to two orders of magnitude cheaper in latency and energy than trained-encoder and generative baselines.

We take OPENER exactly as released and apply it to legal NER, reusing its entity embedder from the Hugging Face Hub unchanged. We evaluate its two operating points: OPENER-ZS, a label-free head (label-name prototypes, transductive refinement, detector fusion), and OPENER-Sup, a balanced linear probe fit on the legal target labels. Our contributions are:
\begin{itemize}
\item A \textbf{transfer study} of a frugal, general-domain open-world NER pipeline to the legal domain, with \emph{no} legal-domain retraining of any component.
\item An evaluation on \textbf{four legal NER corpora spanning three languages} (English, Portuguese and German) along three axes at once (quality (AMI), latency and energy), for both the zero-shot and the supervised typing head, against the same open-world baselines as the general-domain benchmark (GLiNER, GNER, OWNER, a 4-bit LLM).
\item An analysis of \textbf{why the two heads diverge under transfer}. OPENER-ZS remains among the most competitive label-free systems on English legal text, but its standing slips relative to the general domain, where the same head led every compared zero-shot system. OPENER-Sup, in contrast, is barely affected by the domain shift: its balanced linear probe fits separating hyperplanes fine enough to distinguish closely related legal types even where the shared embedding space places their vectors close together, whereas the zero-shot head's label-name matching cannot sharpen that same boundary without supervision. This makes the supervised head the more practical choice whenever legal, or other applied, deployments can supply labelled examples, and points to in-language \emph{detection}, not typing, as the first place further domain adaptation would pay off.
\end{itemize}
