\section{Experiments}
\label{sec:experiments}

\subsection{Datasets}
\label{sec:exp:datasets}
We evaluate on five public legal NER datasets that, together, separate two kinds of distribution shift. \emph{Three are in English}, isolating a \emph{pure domain shift} from the general-domain data the embedder was trained on: \textbf{E-NER} \citep{au2022ener} (annotated U.S.\ SEC/EDGAR filings, $7$ types), \textbf{Indian-Legal} \citep{kalamkar2022indianlegal} (Indian court judgments, $14$ fine-grained legal types), and \textbf{MAPA} \citep{degibert2022mapa} (EUR-Lex legislation, $4$ concrete anonymisation types; English subset). \textbf{LeNER-Br} \citep{luzdearaujo2018lenerbr} (Brazilian court decisions, Portuguese) and \textbf{German-LER} \citep{leitner2020germanler} (German federal court decisions) add a \emph{cross-lingual} shift on top of the domain shift, stressing an English-centric embedder well outside its training distribution. \autoref{tab:legal-datasets} summarises them with the entity-type distribution of their test mentions.

\begin{table*}[t]
\centering
\caption{Legal NER datasets used for the transfer study. ``Sent.'' / ``Ment.'' are the number of test sentences (with $\geq 1$ entity) and test mentions; each entity type is followed by its share (\%) of the test mentions. German-LER is shown with its 7-type \emph{coarse} schema here; its 19-type \emph{fine} schema is used as a specialised-schema stress test in \autoref{sec:exp:analysis}.}
\label{tab:legal-datasets}
\footnotesize
\begin{tabular}{@{}llrrp{0.52\linewidth}@{}}
\toprule
Dataset & Lang & Sent. & Ment. & Entity types (share \% of test mentions) \\
\midrule
E-NER \citep{au2022ener} & en & 877 & 1{,}793 & Business 41.8, Misc 18.3, Legislation/Act 13.5, Location 11.3, Person 7.6, Government 7.1, Court 0.3 \\
Indian-Legal \citep{kalamkar2022indianlegal} & en & 866 & 3{,}028 & Lawyer 16.7, Court 9.3, Respondent 8.8, Other-person 8.7, Provision 8.5, Statute 7.3, Date 7.1, Petitioner 6.0, GPE 5.9, Precedent 5.7, Org 5.1, Judge 5.1, Case-number 3.9, Witness 1.9 \\
MAPA (en) \citep{degibert2022mapa} & en & 72 & 225 & Date 41.3, Person 28.9, Organisation 20.9, Address 8.9 \\
LeNER-Br \citep{luzdearaujo2018lenerbr} & pt & 636 & 1{,}536 & Organização 32.6, Legislação 24.6, Pessoa 15.2, Tempo 12.5, Jurisprudência 12.0, Local 3.1 \\
German-LER \citep{leitner2020germanler} & de & 3{,}018 & 5{,}322 & Legal-norm 38.3, Case-law 23.4, Organization 15.0, Regulation 6.7, Person 6.1, Legal-literature 5.9, Location 4.7 \\
\bottomrule
\end{tabular}
\end{table*}

\subsection{Setup}
\label{sec:exp:setup}
Embedder, detector and heads are exactly those of the general-domain OPENER (\autoref{sec:method}); the entity embedder is loaded from the Hugging Face Hub and is \emph{not} retrained on legal data. Single fixed seed; each test split capped at $1000$ sentences. Since the reused embedder is frozen and no stage of the pipeline is stochastic at inference, the quality scores are deterministic (identical across re-runs and across seeds); following the general-domain protocol, we therefore quantify their sampling uncertainty with a test-set bootstrap ($K{=}1000$ resamples of the gold mentions) rather than a multi-seed \emph{retraining} study, reporting $95\%$ confidence intervals in \autoref{tab:legal-bootstrap}.

\begin{table}[t]
\centering
\caption{Test-set bootstrap of typing-on-gold AMI ($K{=}1000$ resamples of the gold mentions) for the released, frozen embedder: bootstrap mean $\pm$ std with the $95\%$ percentile CI. The point estimates match \autoref{tab:legal-gold}; the narrow $\pm 1.3$--$1.7$ AMI intervals show the reported scores are stable rather than seed-specific.}
\label{tab:legal-bootstrap}
\small
\begin{tabular}{@{}lcc@{}}
\toprule
Dataset & OPENER-Sup AMI & OPENER-ZS AMI \\
\midrule
E-NER (en)      & $70.4 \pm 1.5$~[$67.5$, $73.4$] & $44.4 \pm 1.3$~[$41.7$, $47.0$] \\
LeNER-Br (pt)   & $53.8 \pm 1.7$~[$50.6$, $57.0$] & $32.0 \pm 1.3$~[$29.5$, $34.3$] \\
German-LER (de) & $59.7 \pm 1.5$~[$56.9$, $62.7$] & $23.1 \pm 1.3$~[$20.6$, $25.5$] \\
\bottomrule
\end{tabular}
\end{table}
For end-to-end runs the detector is GLiNER-L at threshold $0.3$, fed the target type names. As \textbf{baselines} we evaluate the same open-world systems as the general-domain benchmark, applied to legal text: GLiNER-S/M/L \cite{zaratiana2024gliner}, the generative GNER-T5-base \cite{ding2024gner}, and a 4-bit Qwen2.5-1.5B-Instruct \cite{qwen2025} prompted for entities -- all zero-shot on the target type names. These are \emph{end-to-end} systems (they detect and type jointly), so they are compared only under the end-to-end protocol, never against typing-on-gold numbers. All quality figures are AMI ($\times 100$); German-LER uses its $7$-type coarse schema in the tables, and its $19$-type fine schema as a stress test below.

\subsection{Results}
\label{sec:exp:results}
Keeping the two protocols apart, \autoref{tab:legal-gold} reports typing-on-gold AMI (OPENER heads, perfect detection) and \autoref{tab:legal-main} the end-to-end comparison against the baselines; \autoref{tab:legal-frugal} gives the accompanying frugality. Three findings stand out.

\textbf{Supervised typing transfers well, including across languages.} Fitting a balanced linear probe on the frozen, general-domain embedder reaches $60.9$ mean typing-on-gold AMI over the five sets -- $70.0$ / $61.9$ / $60.1$ on the three English corpora (E-NER, Indian court judgments, MAPA), and still $53.3$ / $59.1$ on Portuguese LeNER-Br and German German-LER despite the English-centric embedder (\autoref{fig:legal-gold}). The probe adapts to whatever the embedder encodes, so target-label supervision largely absorbs the cross-lingual gap at typing time.

\textbf{Label-free typing is strong in English but degrades cross-lingually.} OPENER-ZS is strong on the three English sets -- $44.0$ (E-NER), $46.3$ (Indian) and its overall best $56.5$ on MAPA, whose concrete type names (\emph{person}, \emph{organisation}, \emph{date}, \emph{address}) are ideal for label-name prototypes -- but drops to $31.6$ (pt) and $22.6$ (de): its English prototypes align well with English legal mentions, far less with Portuguese/German ones. The zero-shot head is where the cross-lingual distribution shift bites hardest.

\textbf{End-to-end, detection sets the ceiling.} On the legal-specific schemas the dedicated systems cluster in a narrow $\approx 33$--$36$ AMI band (\autoref{tab:legal-main}), because the English detectors recover only a fraction of gold mentions (GLiNER-L offset recall $18.7$--$57.4\%$); with so few mentions detected, no typing head can pull far ahead, and OPENER-ZS's detector fusion essentially reverts to the GLiNER-L labels. Where the entity types are concrete and detector-friendly -- MAPA's person / organisation / date / address -- recall rises to $61\%$ and \emph{every} system jumps to $\approx 40$--$52$, GLiNER-M leading ($52.5$); this alone lifts the GLiNER family to the best end-to-end averages ($37.6$--$38.4$). On the other sets the winners are dataset-specific: OWNER \cite{genest2025owner} -- the closest prior work -- on E-NER ($38.4$) and GNER-T5 on German ($39.2$), though OWNER degrades most cross-lingually ($28.4$) and pays an order-of-magnitude higher inference cost (re-encoding and clustering every mention), and Qwen trails throughout ($24.7$ mean) while being the slowest. No system dominates end-to-end -- the ceiling is detection, not typing -- and OPENER-Sup/ZS sit mid-band ($34.1$/$35.0$ average) at a fraction of the cost, consistent with the typing-on-gold results where OPENER-Sup leads by a wide margin.

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar, bar width=8pt,
  width=\linewidth, height=5cm,
  ymin=0, ymax=82,
  ylabel={Typing-on-gold AMI},
  symbolic x coords={E-NER (en), Indian (en), MAPA (en), LeNER-Br (pt), German (de)},
  xtick=data, x tick label style={font=\footnotesize, rotate=20, anchor=east},
  legend style={at={(0.5,1.03)}, anchor=south, legend columns=2, font=\small, draw=none},
  ymajorgrids, grid style={gray!20},
  nodes near coords, nodes near coords style={font=\tiny},
  enlarge x limits=0.16,
]
\addplot[fill=yellow!75!orange, draw=black!40] coordinates {(E-NER (en),44.0) (Indian (en),46.3) (MAPA (en),56.5) (LeNER-Br (pt),31.6) (German (de),22.6)};
\addplot[fill=red!55, draw=black!40] coordinates {(E-NER (en),70.0) (Indian (en),61.9) (MAPA (en),60.1) (LeNER-Br (pt),53.3) (German (de),59.1)};
\legend{OPENER-ZS, OPENER-Sup}
\end{axis}
\end{tikzpicture}
\caption{Typing-on-gold AMI by dataset (three English, then Portuguese and German). The supervised probe (OPENER-Sup) stays high everywhere; the label-free head (OPENER-ZS) is strong on the three English sets -- best of all on MAPA's concrete types ($56.5$) -- but degrades sharply under the cross-lingual shift.}
\label{fig:legal-gold}
\end{figure}

\begin{table}[t]
\centering
\caption{\textbf{Typing-on-gold} AMI ($\times100$, $\uparrow$): entity typing on gold mentions (perfect detection), which isolates the typing head. German-LER uses its $7$-type coarse schema (the $19$-type fine schema is discussed in \autoref{sec:exp:analysis}). Bold marks the best per column.}
\label{tab:legal-gold}
\small
\begin{tabular}{@{}lcccccc@{}}
\toprule
System & E-NER & Indian & MAPA & LeNER-Br & German & Mean \\
       & (en)  & (en)   & (en) & (pt)     & (de)   &      \\
\midrule
\rowcolor{hlyellow} OPENER-ZS & $44.0$ & $46.3$ & $56.5$ & $31.6$ & $22.6$ & $40.2$ \\
\rowcolor{hlred} OPENER-Sup   & $\mathbf{70.0}$ & $\mathbf{61.9}$ & $\mathbf{60.1}$ & $\mathbf{53.3}$ & $\mathbf{59.1}$ & $\mathbf{60.9}$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{table*}[t]
\centering
\caption{Open-world legal NER, \textbf{end-to-end} AMI ($\times100$, $\uparrow$): mentions produced by the detector, with offset/sentinel alignment (the protocol comparable across all systems). Systems grouped by supervision; GLiNER/GNER/Qwen are fed the target type names zero-shot, OWNER \cite{genest2025owner} is transferred from a CoNLL-2003-trained model (no target labels), OPENER-Sup fits a balanced probe on the target labels. Bold marks the best per column. $^\ddagger$Qwen2.5-1.5B on a $200$-sentence subsample (indicative). German-LER: $7$-type coarse schema. Typing-on-gold is reported separately in \autoref{tab:legal-gold}.}
\label{tab:legal-main}
\small
\begin{tabular}{@{}lcccccc@{}}
\toprule
Model & E-NER (en) & Indian (en) & MAPA (en) & LeNER-Br (pt) & German (de) & Avg \\
\midrule
\multicolumn{7}{l}{\emph{Zero-shot (no target labels)}} \\
GLiNER-S & $33.4$ & $33.9$ & $49.4$ & $33.7$ & $31.1$ & $36.3$ \\
GLiNER-M & $33.9$ & $34.8$ & $\mathbf{52.5}$ & $34.8$ & $36.1$ & $\mathbf{38.4}$ \\
GLiNER-L & $34.7$ & $\mathbf{35.2}$ & $47.8$ & $\mathbf{36.3}$ & $34.0$ & $37.6$ \\
GNER-T5  & $28.7$ & $28.6$ & $38.3$ & $33.2$ & $\mathbf{39.2}$ & $33.6$ \\
Qwen2.5-1.5B$^\ddagger$ & $24.9$ & $28.1$ & $31.1$ & $24.9$ & $14.5$ & $24.7$ \\
OWNER    & $\mathbf{38.4}$ & $33.3$ & $38.4$ & $34.8$ & $28.4$ & $34.7$ \\
\rowcolor{hlyellow} OPENER-ZS & $33.2$ & $34.0$ & $39.8$ & $34.6$ & $33.2$ & $35.0$ \\
\midrule
\multicolumn{7}{l}{\emph{Supervised (target labels)}} \\
\rowcolor{hlred} OPENER-Sup & $35.8$ & $32.8$ & $35.4$ & $32.4$ & $34.2$ & $34.1$ \\
\bottomrule
\end{tabular}
\end{table*}

\begin{table}[t]
\centering
\caption{Efficiency of the end-to-end systems (median over the five datasets): median per-sentence latency and energy per run, alongside end-to-end AMI. OPENER stays in the leading accuracy band while remaining one to two orders of magnitude cheaper than the generative systems. $^\ddagger$Qwen on the $200$-sentence subsample. $^\S$OWNER: wall-clock (includes per-dataset model load and mention clustering), TDP estimate.}
\label{tab:legal-frugal}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
System & AMI & Latency p50 (ms) & Energy (Wh) \\
\midrule
GLiNER-S & $36.3$ & $\mathbf{22}$ & $\mathbf{0.4}$ \\
GLiNER-M & $\mathbf{38.4}$ & $36$ & $0.6$ \\
GLiNER-L & $37.6$ & $78$ & $1.2$ \\
GNER-T5  & $33.6$ & $7053$ & $67$ \\
Qwen2.5-1.5B$^\ddagger$ & $24.7$ & $4997$ & $16$ \\
OWNER$^\S$ & $34.7$ & $420$ & $11$ \\
\rowcolor{hlyellow} OPENER-ZS & $35.0$ & $115$ & $2.1$ \\
\rowcolor{hlred} OPENER-Sup & $34.1$ & $123$ & $2.1$ \\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar, bar width=10pt,
  width=\linewidth, height=5cm,
  ymode=log, ymin=10, ymax=20000, log origin=infty,
  ylabel={Latency p50 (ms/sent, log)},
  symbolic x coords={GLiNER-S,GLiNER-M,GLiNER-L,OWNER,OP-ZS,OP-Sup,Qwen,GNER},
  xtick=data, x tick label style={rotate=35, anchor=east, font=\footnotesize},
  ymajorgrids, grid style={gray!20}, enlarge x limits=0.09,
]
\addplot[fill=blue!30, draw=black!40] coordinates {(GLiNER-S,22)(GLiNER-M,36)(GLiNER-L,78)(OWNER,420)(Qwen,4997)(GNER,7053)};
\addplot[fill=red!55, draw=black!40] coordinates {(OP-ZS,115)(OP-Sup,123)};
\legend{baselines, OPENER}
\end{axis}
\end{tikzpicture}
\caption{End-to-end per-sentence latency (log scale, median over the five datasets). OPENER (red) and the GLiNER family run in $22$--$123$\,ms, whereas the generative GNER-T5 and Qwen are $40$--$60\times$ slower ($\approx 5$--$7$\,s per sentence).}
\label{fig:legal-efficiency}
\end{figure}

The three-axis picture (\autoref{tab:legal-frugal}, \autoref{fig:legal-efficiency}) mirrors the general-domain study: OPENER types a sentence end-to-end in ${\approx}120$\,ms for ${\approx}2.1$\,Wh -- on par with the GLiNER family and OWNER, and \emph{one to two orders of magnitude} below the generative GNER-T5 ($5.7$\,s, $67$\,Wh) and Qwen ($4.9$\,s, $16$\,Wh). Zero-shot typing on gold is cheaper still ($0.5$\,Wh, $32$\,ms), as it embeds no training split. OPENER thus keeps the leading accuracy band at a small, fixed inference cost, which is the operating point that makes it deployable on modest hardware.

\subsection{Analysis}
\label{sec:exp:analysis}
Two distribution shifts act together here, and the results separate them cleanly. On \textbf{English} legal text (E-NER), where only the \emph{domain} shifts, transfer is strong: OPENER-Sup reaches $70.0$ gold AMI and OPENER-ZS a usable $44.0$ with no target labels at all, confirming that the general-domain contrastive embedder already places legal entity types (business, court, government body, legislation) in largely separable regions. Adding a \textbf{cross-lingual} shift (Portuguese, German) widens the gap between the two heads: the supervised probe stays high ($53.3$/$59.1$) because it re-fits on target labels, whereas the label-free head drops to $22$--$32$ because its English prototypes and the English embedder no longer match the target-language mentions.

The \textbf{specialised-schema} stress test points the same way. On German-LER's fine $19$-type schema, OPENER-Sup still reaches $62.0$ gold AMI -- the embedder separates even fine legal distinctions once a probe names them -- but OPENER-ZS falls to $22.1$, as many fine German type codes have no faithful English label-name prototype.

The binding constraint end-to-end is \textbf{detection}, not typing: GLiNER-L's offset recall falls from $57.4\%$ (en) to $18.7\%$ (de), so on the legal-specific schemas the end-to-end scores compress to ${\approx}34$--$35$ regardless of the typing head. This localises where domain adaptation would pay off first: a light, in-language adaptation of the \emph{detector} (or a multilingual one) would lift the ceiling far more than a better typing head. Within our frugal, retraining-free budget, OPENER-Sup is the safe operating point across languages, while OPENER-ZS is attractive only when the target language is close to the embedder's (English) training distribution.

\textbf{A controlled language ablation} attributes the cross-lingual cost to the pipeline, not to OPENER. The MAPA corpus (identical annotation scheme, concrete types) exists in both English and French, so we re-run every system on its French subset ($156$ test sentences). Switching English$\to$French is costly for OPENER's typing -- $60.1\!\to\!44.2$ gold AMI for OPENER-Sup ($-15.9$) and $56.5\!\to\!36.3$ for OPENER-ZS -- but this is \emph{not} specific to OPENER: end-to-end the English GLiNER-M loses $11.9$ AMI ($52.5\!\to\!40.6$) and OWNER $5.0$, whereas the multilingual Qwen2.5-1.5B is essentially unaffected ($+0.8$, \autoref{tab:legal-mapafr}). The whole English-centric stack -- embedder \emph{and} detector -- degrades together on a new language, so a multilingual backbone or a light in-language adaptation, rather than a better typing head, is what would recover the gap.

\begin{table}[t]
\centering
\caption{Controlled language ablation on the MAPA corpus (same annotation scheme, English vs French subset). End-to-end AMI ($\times100$, $\uparrow$): the cross-lingual shift degrades the whole English-centric pipeline; only the multilingual Qwen is unaffected.}
\label{tab:legal-mapafr}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
System (end-to-end) & MAPA-en & MAPA-fr & $\Delta$ \\
\midrule
GLiNER-M & $52.5$ & $40.6$ & $-11.9$ \\
GNER-T5  & $38.3$ & $31.4$ & $-6.9$ \\
OWNER    & $38.4$ & $33.4$ & $-5.0$ \\
Qwen2.5-1.5B & $31.1$ & $31.9$ & $\mathbf{+0.8}$ \\
\rowcolor{hlyellow} OPENER-ZS & $39.8$ & $35.7$ & $-4.1$ \\
\rowcolor{hlred} OPENER-Sup & $35.4$ & $31.9$ & $-3.5$ \\
\bottomrule
\end{tabular}
\end{table}
