\section{Conclusion}
\label{sec:conclusion}

We studied whether OPENER, a frugal open-world NER pipeline built from off-the-shelf pretrained parts and trained only on general-domain data, transfers to legal NER without any legal-domain retraining, evaluating a label-free head (OPENER-ZS: prototypes, transductive refinement, detector fusion) and a supervised linear probe (OPENER-Sup) on four legal NER corpora spanning three languages (English, Portuguese and German) against the same open-world baselines as the general-domain benchmark.

The two heads diverge sharply under transfer. OPENER-ZS remains a usable, competitive label-free option on English legal text, but its standing \emph{slips relative to the general domain}, where the same head led every compared zero-shot system end-to-end: here it trails the GLiNER family and OWNER on average, and its gold AMI collapses cross-lingually as its English-centric prototypes stop matching target-language mentions. OPENER-Sup, in contrast, is almost unaffected by the domain shift, its gold AMI within a point of its general-domain score even on German-LER's fine $19$-type schema, and it posts the best end-to-end AMI throughout, a statistically significant lead. The reason is geometric rather than purely statistical: a balanced linear probe fits separating hyperplanes precise enough to pull apart legal types whose vectors sit close together in a general-purpose embedding space (a lawyer from a judge, a statute from a case citation), a distinction the zero-shot head's single label-name match cannot sharpen without supervision. This has a direct practical reading: whenever legal, or other applied, deployments can supply even a modest labelled sample, the supervised head is the safer and more accurate operating point, and OPENER-ZS is best reserved for domains and languages close to the embedder's own training distribution.

Echoing the general-domain study, end-to-end quality is still bounded first by mention \emph{detection}, regardless of the typing head: light, in-language adaptation of the detector, rather than of the embedder, is therefore the natural next step, keeping the frugal, decoupled detect-then-type design intact.

\section*{Limitations}
Test splits are capped at $1000$ sentences. OPENER scores are means over three embedder-retraining seeds: the supervised and end-to-end scores are stable (std ${\leq}1.4$ AMI), but the label-free prototypes vary by up to ${\pm}6$ AMI across retrainings, so single-seed zero-shot numbers should be read with care. Inference is deterministic for a fixed embedder, and test-set bootstraps on the released model confirm both the typing-on-gold scores (std ${\pm}0.8$--$1.7$ AMI) and the significance of OPENER-Sup's end-to-end lead, stable across resampling seeds $42/7/123$. More fundamentally, both the embedder and the detector are English-centric and reused without any legal- or language-domain adaptation; the detector's low cross-lingual recall (e.g.\ $18.7\%$ on German) caps end-to-end quality and is the first thing a follow-up should address.

\section*{Ethics Statement}
This work runs and evaluates entity recognition on four publicly released legal NER benchmarks (E-NER over SEC/EDGAR filings, an Indian court-judgment set, LeNER-Br, and German-LER), used only for evaluation and under their original licenses. Legal documents contain personal data such as the names of parties, judges and counsel. We introduce no new annotations and release no data, we only run inference over the existing public test splits, and we report no individual-level content, so the study adds no privacy exposure and makes no attempt at re-identification.

Open-world NER is error-prone, and our own results show recall dropping sharply outside English (down to $18.7\%$ on German). The pipeline is therefore meant as an assistive tool that surfaces candidate entities for human review, not as a component of automated legal decision-making, where a missed or mistyped entity can carry material consequences. A missed entity is silent, so a human in the loop remains necessary, especially cross-lingually. Finally, a central motivation of this study is energy frugality. We report per-sentence latency and energy alongside accuracy precisely to make the environmental cost of deployment explicit, and to argue for models that are cheap enough to run on modest hardware.
